---
title: "p8015_hw3_yx2507"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
devtools::install_github("thomasp85/patchwork")
library(tidyverse)
library(patchwork)
library(ggridges)
library(lubridate)
```
# Problem 1
load the dataset
```{r}
library(p8105.datasets)
data("instacart")
```
decription of the dataset
## 1 How many aisles are there, and which aisles are the most items ordered from?
There are `r nrow(instacart)` rows and `r ncol(instacart)` columns. 
there are `r count(instacart, vars= "aisle")` 
```{r}
instacart%>%
  distinct(aisle)%>%
  nrow()

instacart%>% 
  count(aisle)%>%
  arrange(desc(n))
```
There are 134 distinct aisles in the dataset and the most frequently ordered is fresh vegetables which is ordered by 150609 times.
## 1.2 Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r, plot1, fig.width=20, fig.height=8}
instacart%>%
  count(aisle)%>%
  mutate(aisle = fct_reorder(aisle, desc(n)))%>%
  filter(n>10000)%>%
  ggplot(aes(x = aisle, y = n))+
   scale_y_discrete(breaks = c(200000,20000, 2000, 200),labels=c("200000","20000", "2000", "200"), limits = c(10000,200000)) +
  geom_point(aes(color = aisle))+
  labs(
      title = "Number of items ordered in each aisle",
      x = "Aisle",
      y = "Number of items"
    ) +
theme(axis.text.x = element_text(angle = 70, hjust = 1))
```
## 1.3 Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>% 
  group_by(aisle, product_name)  %>% 
  count() %>% 
  group_by(aisle) %>%
  mutate(order_rank=min_rank(desc(n)))%>%
  filter(order_rank < 4)%>%
  select(aisle, product_name)%>%
   knitr::kable(
    col.names = c("Aisle", "Most popular item"), 
    format = 'html', 
    caption = "Table: The most popular item in three aisles"
  )
```

## 1.4 Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
 instacart %>% 
  mutate(order_dow = factor(order_dow, labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))) %>% 
  filter(product_name %in% c('Pink Lady Apples', "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow)%>%
  summarise(mean=mean(order_hour_of_day))%>%
  select(product_name,order_dow,mean)%>%
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean"
  ) %>% 
knitr::kable( digits = 1，
  col.names = c("Product Name", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"),
  caption = "Table of mean order hours for pink lady apples and coffee ice creams in each day ",
  format = 'html'
)
  
  
```
# prolem 2
load the dataset

```{r}

library(p8105.datasets)
data("brfss_smart2010")

```
## 2.1
First, do some data cleaning:
format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”
```{r}
health_data=
  brfss_smart2010%>%
  janitor::clean_names()%>%
  filter(topic %in% "Overall Health" )%>%
   mutate(response = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor")))

view(health_data)
  

```

In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
health_data%>%
  filter(year %in% "2002")%>%
  group_by(locationabbr)%>%
  count()%>%
  filter(n>7)%>%
  knitr::kable(
  )

health_data%>%
  filter(year %in% "2010")%>%
  group_by(locationabbr) %>%
  count()%>%
  filter(n>7)%>%
  knitr::kable()

```
Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
```{r}


 health_subdata = 
  health_data%>%
  filter(response %in% "Excellent")%>%
  group_by(year, locationabbr)%>%
  summarize(mean_datavalue=mean(data_value, na.rm = TRUE))%>%
  select(year, locationabbr, mean_datavalue)
  
  ## make a plot
plot_average = 
  health_subdata%>%
ggplot(aes(x=year, y=mean_datavalue, color = locationabbr))+
  geom_line(alpha = .5)



```
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State
```{r}
plot_data=
health_data%>%
  filter(response %in% c("Poor", "Excellent"), year %in% c("2006","2010"), locationabbr == "NY")%>%
  select(year, response, locationabbr, data_value)%>%
    ggplot(aes(x=response, y=data_value, Color=response))+
  geom_point()+
  facet_grid(. ~ year)
    
  
```

