---
title: "p8015_hw3_yx2507"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
devtools::install_github("thomasp85/patchwork")
library(tidyverse)
library(patchwork)
library(ggridges)
library(lubridate)
```
# Problem 1
load the dataset
```{r}
library(p8105.datasets)
data("instacart")
```
decription of the dataset
## 1 How many aisles are there, and which aisles are the most items ordered from?
There are `r nrow(instacart)` rows and `r ncol(instacart)` columns. 
there are `r count(instacart, vars= "aisle")` 
```{r}
instacart%>%
  distinct(aisle)%>%
  nrow()

instacart%>% 
  count(aisle)%>%
  arrange(desc(n))
```
There are 134 distinct aisles in the dataset and the most frequently ordered is fresh vegetables which is ordered by 150609 times.
## 1.2 Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r, plot1, fig.width=20, fig.height=8}
instacart%>%
  count(aisle)%>%
  mutate(aisle = fct_reorder(aisle, desc(n)))%>%
  ggplot(aes(x = aisle, y = n))+
   scale_y_discrete(breaks = c(200000,20000, 2000, 200),labels=c("200000","20000", "2000", "200"), limits = c(10000,200000)) +
  geom_point()+
  labs(
      title = "Number of items ordered in each aisle",
      x = "Aisle",
      y = "Number of items"
    )
```
## 1.3 Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>% 
  group_by(aisle, product_name)  %>% 
  count() %>% 
  group_by(aisle) %>% 
  filter(n==max(n))%>%
  select(aisle, product_name)%>%
   knitr::kable(
    col.names = c("Aisle", "Most popular item"), 
    format = 'html', 
    caption = "Table: The most popular item in three aisles"
  )
```

